Active Learning for SAT Solver Benchmarking

### 2022-10-10 paper/paper.tex
introduced \hat big-A

Let benchmark instances $\mathcal{I}$, SAT solvers $\mathcal{A}$, and runtimes $r : \mathcal{A} \times \mathcal{I} \rightarrow \left[0, \tau\right]$ be given. 
At the core of the model $\mathcal{M}$ is a prediction function $f : \left(\mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace\right) \times \mathcal{I} \rightarrow \mathbb{R}$ that powers the decisions within the previously mentioned three steps.

To be applicable to $\hat{a}$, we extend \emph{score} by replacing $\gamma\!\left(a, i\right)$ by the model's predictions $f\!: \left(\mathcal{A} \cup 
\left\lbrace \hat{a} \right\rbrace\right) \times \mathcal{I} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$.

@@

Selecting an instance based on the model is the core functionality of our method (cf. Algorithm~\ref{algALBenchmark}, Line 2). 
In this section, we introduce our sampling strategies that make use of the model's label-prediction function $f$ which are inspired by existing work within the realms of active learning~\cite{settles2009active}.
We implement a model-uncertainty based approach and a model-based information-gain sampling strategy.
%, and a voting-based committee-disagreement approach.

These methods require the model's predictions to also include probabilities for the $k$ possible runtime labels, i.e., 
\begin{equation*}
  f'\!: \left(\mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace\right) \times \mathcal{I} \rightarrow \left[0, 1\right]^k
\end{equation*}

@@

@@

\subsubsection{Wilcoxon Stopping Criterion}
The Wilcoxon stopping criterion stops the iterative active-learning algorithm when we are on average certain enough that the predicted runtime labels are sufficiently different in comparison to the other solvers.
To be more specific, we stop when
\begin{equation}
  W_{\hat{a}} := \frac{1}{\left\lvert \mathcal{A} \right\rvert} \sum_{a \in \mathcal{A}} \operatorname{wilcoxon}_{\alpha}\!\left(\gamma\!\left(a, \mathcal{I}\right), f\!\left(\hat{a}, \mathcal{I}\right) \right)
\end{equation}
drops below a certain threshold.
%JB: if this is info gain regarding runtimes, doesn't Eq. (4) mean that we would prefer instances with low entropy of runtime distribution after considering the new solver, i.e., instances that are not very discriminatory?
%TF: I don't think so. We want to maximize the reduction in entropy of the runtime labels. Low entropy afterward -> we know that the solver behaves similarly to existing ones.
Thereby, \emph{wilcoxon} denotes the $p$-value of a Wilcoxon signed-rank test of the runtime-label distributions and $\gamma\!\left(a, \mathcal{I}\right)$ and $f\!\left(\hat{a}, \mathcal{I}\right)$ are the natural set extensions on $\gamma$ and $f$ respectively.
To improve the stability of this criterion, we use an exponential moving average

\begin{equation}
  W_{\exp}^{\left(m\right)} := \beta W_{\hat{a}} + \left(1 - \beta\right) W_{\exp}^{\left(m - 1\right)} \enspace \textrm{,}
\end{equation}

whereby $m$ denotes the count of iterations within Algorithm~\ref{algALBenchmark} and $W_{\exp}^{\left(0\right)} := 1$.

>>>>

Let benchmark instances $\mathcal{I}$, SAT solvers $\mathcal{A}$, and runtimes $r : \mathcal{A} \times \mathcal{I} \rightarrow \left[0, \tau\right]$ be given. 
We denote the new solver to be ranked by $\hat a \not\in A$ and define $\hat A := A \cup \hat a$ to denote all solvers including the new solver. 
At the core of the model $\mathcal{M}$ is a prediction function $f : \mathcal{\hat A} \times \mathcal{I} \rightarrow \mathbb{R}$ that powers the decisions within the previously mentioned three steps.

To be applicable to $\hat{a}$, we extend \emph{score} by replacing $\gamma(a, e)$ by the model's predictions $f : \mathcal{\hat A} \times \mathcal{I} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$.

@@

Selecting an instance based on the model is the core functionality of our method (cf. Algorithm~\ref{algALBenchmark}, Line 2). 
In this section, we introduce our sampling strategies that make use of the model's label-prediction function $f$ which are inspired by existing work within the realms of active learning~\cite{settles2009active}.
We implement a model-uncertainty based approach and a model-based information-gain sampling strategy.

These methods require the model's predictions to also include probabilities for the $k$ possible runtime labels, i.e., 
\begin{equation*}
  f' : \mathcal{\hat A} \times \mathcal{I} \rightarrow \left[0, 1\right]^k
\end{equation*}

@@

%JB: if this is info gain regarding runtimes, doesn't Eq. (4) mean that we would prefer instances with low entropy of runtime distribution after considering the new solver, i.e., instances that are not very discriminatory?
%TF: I don't think so. We want to maximize the reduction in entropy of the runtime labels. Low entropy afterward -> we know that the solver behaves similarly to existing ones.

@@

\subsubsection{Wilcoxon Stopping Criterion}
The Wilcoxon stopping criterion stops the iterative active learning algorithm when we are on average certain enough that the predicted runtime labels are sufficiently different in comparison to the other solvers.

To be more specific, we stop when
\begin{equation}
  W_{\hat{a}} := \frac{1}{\left\lvert \mathcal{A} \right\rvert} \sum_{a \in \mathcal{A}} \operatorname{wilcoxon}_{\alpha}\!\left(\gamma\!\left(a, \mathcal{I}\right), f\!\left(\hat{a}, \mathcal{I}\right) \right)
\end{equation}
drops below a certain threshold.
Thereby, \emph{wilcoxon} denotes the $p$-value of a Wilcoxon signed-rank test of the runtime-label distributions and $\gamma\!\left(a, \mathcal{I}\right)$ and $f\!\left(\hat{a}, \mathcal{I}\right)$ are the natural set extensions on $\gamma$ and $f$ respectively.
To improve the stability of this criterion, we use an exponential moving average to smooth out outliers.

%whereby $m$ denotes the count of iterations within Algorithm~\ref{algALBenchmark} and $W_{\exp}^{\left(0\right)} := 1$.
%\begin{equation}
%  W_{\exp}^{\left(m\right)} := \beta W_{\hat{a}} + \left(1 - \beta\right) W_{\exp}^{\left(m - 1\right)}
%\end{equation}
